---
title: "R Notebook"
output: html_notebook
---

Import the Training and Testing Data Sets:

```{r}
library(readr)
train <- read_csv("C:/Users/anant/Desktop/ML Project/train.csv")
test <- read_csv("C:/Users/anant/Desktop/ML Project/test.csv")
```

Coverting the label column to the type factor:

```{r}
train$label  <- as.factor(train$label)

```


Exploratory Data Analysis:

Number of Entries and Features in the Training Sets:

```{r}
dim(train)

```

Number of Entries and Features in the Testing Set:

```{r}
dim(test)

```

Let us now visualize the number of each digit in the training set:

```{r}
p1 <- ggplot(train, aes(x =label, color = label))+ geom_bar(stat = "Count", fill = "white") 
p1

```

But it would be great to know the actual numbers too:

The Following Table Distribution of digits in the Training Set:

```{r}
cat("Distribution in Training Set:\n")
table(train$label)

```

A good starting Point before we develop the models would be to visualize the digits in the datase:

```{r}

digits_vis <- sample(1:nrow(train),25)


digits<- t(train[digits_vis,-1])
digits_matrix <- lapply(1:25,function(x) matrix(digits[,x],ncol=28))


par(mfrow=c(5,5),mar=c(.1,.1,.1,.1))

for(i in 1:25) 
{
  for(j in 1:28) 
    {
    digits_matrix[[i]][j,] <- rev(digits_matrix[[i]][j,])
    }
  
  image(digits_matrix[[i]],col=grey.colors(225), axes = FALSE)
}

```

Splitting the Dataset into Training and Testing Sets:
```{r}
pd <- sample(2,nrow(train),replace = TRUE, prob=c(0.8,0.2))
train_df <-  train[pd==1,]
test_df  <-  train[pd==2,]

```

Data Preproccessing:

```{r}
library(caret)
nzr <- nearZeroVar(train_df[,-1],saveMetrics=T,freqCut=10000/1,uniqueCut=1/7)
sum(nzr$zeroVar)

```

Total Number :

```{r}
sum(nzr$nzv)

```

```{r}
cutvar <- rownames(nzr[nzr$nzv==TRUE,])
var <- setdiff(names(train_df),cutvar)
train_df <- train_df[,var]

```

PCA with Data : 

```{r}
label <- as.factor(train_df[[1]])
train_df$label <- NULL
train_df <- train_df/255
covtrain <- cov(train_df)

```

Number of Components Representaive of the Dataset:

```{r}
train_pc <- prcomp(covtrain)
varex <- train_pc$sdev^2/sum(train_pc$sdev^2)
varcum <- cumsum(varex)
result <- data.frame(num=1:length(train_pc$sdev),
                         ex=varex,
                         cum=varcum)

plot(result$num,result$cum,type="b",xlim=c(0,100),
     main="Variance Explained by Top 100 Components",
     xlab="Number of Components",ylab="Variance Explained")
abline(v=25,lty=2)

```

Plotting the Principal Components :

```{r}
train_score <- as.matrix(train_df) %*% train_pc$rotation[,1:25]
train_df <- cbind(label,as.data.frame(train_score))

colors <- rainbow(length(unique(train_df$label)))
names(colors) <- unique(train_df$label)

plot(train_df$PC1,train_df$PC2,type="n",main="First Two Principal Components")
text(train_df$PC1,train_df$PC2,label=train_df$label,col=colors[train_df$label])
```


Random Forest:

```{r}
library(randomForest)
rf_model <- randomForest(label~.,data = train_df, ntree = 100)
summary(rf_model)
```

Prediction Based on Random Forests:

```{r}
newtest1 <- test_df[,-1]
newtest1 <- newtest1[,var[-1]]/255
newtest1 <- as.matrix(newtest1) %*% train_pc$rotation[,1:25]
newtest1 <- as.data.frame(newtest1)
pred <- predict(rf_model,newtest1)
```

Confusion Matrix:

```{r}
table1 <- table(pred,test_df$label)
print(table1)
```

Accuracy: 

```{r}
Accuracy <- sum(diag(table1))/nrow(test_df)
Accuracy

```

K-Nearest neigbbours:

```{r}
library(class)
model.knn <- knn(train = train_df[,-1], test = newtest1,cl = train_df$label, k=7)

table2 <- table(`Actual Class` = test_df$label, `Predicted Class` = model.knn)

```

Confusion Matrix for KNN:

```{r}
print(table2)
Accuracy2 <- sum(diag(table2))/nrow(test_df)
Accuracy2
```

Descision Tree:

```{r}
model.rpart <- rpart(label~., method = "class", data = train_df)

```

Printing the Tree:

```{r}
plot(model.rpart, uniform = TRUE, main = "Classification (RPART). Tree of Handwritten Digit Recognition ")
text(model.rpart, all = TRUE, cex = 0.75)

```

Predictions:

```{r}
prediction.rpart <- predict(model.rpart, newdata = test_df, type = "class")
table3 <-table(`Actual Class` = test_df$label, `Predicted Class` = prediction.rpart)
print(table3)

```

Accuracy:

```{r}
Accuracy3 <- sum(diag(table3))/nrow(test_df)
Accuracy3

```


